---
title: "DSC365: Introduction to Data Science"
author: "Web-Scraping"
date: "April 23, 2026"
format: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
```

## Resources

- "Web Scraping in R: rvest Tutorial": https://www.datacamp.com/community/tutorials/r-web-scraping-rvest
- "rvest: easy web scraping with R": https://blog.rstudio.com/2014/11/24/rvest-easy-web-scraping-with-r/
- "Web scraping tutorial in R": https://towardsdatascience.com/web-scraping-tutorial-in-r-5e71fd107f32
- From the person that helped write the package: https://r4ds.hadley.nz/webscraping.html

## Web scraping?

__Web scraping__ is the process of extracting information from websites automatically and transforming it into a structured data set.

- There's TONS of data freely available online, but...
- ...the data is provided in an unstructured format: you can always copy & paste, but it's time-consuming and prone to errors.

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
```

## What is webscraping? 

- Extract data from websites 
    + Tables
    + Links to other websites
    + Text
    
Two different scenarios:

1. Screen scraping 
2. Web APIs (application programming interface) - use when possible but out of the scope of this class.


## Why webscrape? 

- Because copy-paste is awful 
- Because it's fast
- Because you can automate it

## Before scraping, do some googling!

- If resource is well-known (e.g. Twitter, Fitbit, etc.), there is *probably* an existing R package for it.
- [ropensci](https://ropensci.org/) has a [ton of R packages](https://ropensci.org/packages/) providing easy-to-use interfaces to open data.
- The [Web Technologies and Services CRAN Task View](http://cran.r-project.org/web/views/WebTechnologies.html) is a great overview of various tools for working with data that lives on the web in R.

## Scraping ethics and legalities

Check to make sure you're allowed! Weather.com allows users to access the data...

```{r warning=FALSE}
# install.packages("robotstxt")
library(robotstxt)
paths_allowed("https://weather.com")
```

... while other sites like Twitter (now) don't.

```{r warning=FALSE}
paths_allowed("https://x.com")
```

Another way to do that is just append “/robots.txt” to the end of the URL of the website you are targeting, which will also tell you whether the website is allowed to be scraped. 

**General Principle:** If the data is public, non-personal, and factual, you're likely to be ok. 
  + Connected to the sites' terms and conditions, personally identificaible information, and copyright
  + If the data isn’t public, non-personal, or factual *or* you’re scraping the data specifically to make money with it, you’ll need to talk to a lawyer


## A web of *messy* data!

- Recall the concept of [tidy data](http://vita.had.co.nz/papers/tidy-data.pdf).
- Data is in a table where
    * 1 row == 1 observation
    * 1 column == 1 variable (observational attribute)
- Parsing web data (HTML/XML/JSON) is easy (for computers)
- Getting it in a tidy form is typically *not easy*.
- Knowing a bit about modern tools & web technologies makes it *much* easier.


## Hypertext Markup Language (HTML)

Most of the data on the web is still largely available as HTML - while it is structured (hierarchical/tree based) it often is not available in a form useful for analysis (flat / tidy).

```html
<html>
  <head>
    <title>This is a title</title>
  </head>
  <body>
    <p align="center">Hello world!</p>
  </body>
</html>
```

HTML has a hierarchical structure formed by elements which consist of a start tag (e.g., <tag>), optional attributes (id='first'), an end tag4 (like </tag>), and contents (everything in between the start and end tag).

Web scraping is possible because most pages that contain data that you want to scrape generally have a consistent structure.
    

## `rvest`

`rvest` is a package that makes basic processing and manipulation of HTML data straight forward.

```{r}
#install.packages('rvest')
library(rvest)
```

`rvest` core functions:

- `read_html` - read HTML data from a url or character string.
- `html_nodes` - select specified nodes from the HTML document using CSS selectors.
  + extracts all elements from the page `x` that have the tag / class / id `path`. (Use SelectorGadget to determine `path`.)
- `html_table` - parse an HTML table into a data frame.
- `html_text` - extract tag pairs' content.
  + extracts all text from the nodeset `x` 
- `html_name` - extract tags' names.
- `html_attrs` - extract all of each tag's attributes.
- `html_attr` - extract tags' attribute value by name.


## SelectorGadget

+ SelectorGadget is a javascript bookmark let to determine the css selectors of pieces of a website we want to extract.
+ Bookmark the SelectorGadget link, https://rvest.tidyverse.org/articles/selectorgadget.html, then click on it to use it (or add the chrome extension)
+ When SelectorGadget is active, pieces of the website are highlighted in orange/green/red.

### Star Wars Example

Let's start with a simple example

1. Open: https://rvest.tidyverse.org/articles/starwars.html

2. Tell R what web-site we want to scrape

```{r}
html <- read_html("https://rvest.tidyverse.org/articles/starwars.html")
```

3. Our goal is to turn this data into a 7 row data frame with variables title, year, director, and plot summary

**Method 1**: Get Elements Individually

```{r}
title <- html %>% html_nodes("#main h2") %>% html_text2()
director <-  html %>% html_nodes(".director") %>% html_text2()
plot_summary <- html %>% html_nodes(".crawl") %>% html_text2()

star_wars <- tibble(
  title = title,
  director = director,
  plot_summary = plot_summary
)
```


**Method 2**: Pull Entire Section at once.

```{r}
section <- html %>% html_elements("section")

star_wars2<- tibble(
  title = section %>% 
    html_element("h2") %>% 
    html_text2(),
  released = section %>% 
    html_element("p") %>% 
    html_text2() %>% 
    str_remove("Released: ") %>%
    parse_date(),
  director = section %>% 
    html_element(".director") %>% 
    html_text2(),
  intro = section %>% 
    html_element(".crawl") %>% 
    html_text2()
)

```


## Your Turn

1. Use SelectorGadget on https://www.baseball-reference.com/teams/KCR/2026-roster.shtml.
2. Pull from "Current 40 Man Roster" Table Only
  + Player's Name
  + Player's Age
  + Player's Position
  + Player's first year in MLB (`1stYr`)

```{r}
url <- "https://www.baseball-reference.com/teams/KCR/2026-roster.shtml"
html <- read_html(url)

name <- html %>% html_nodes("#the40man a") %>% html_text()
position <- html %>% html_nodes("#the40man .right:nth-child(5)") %>% html_text()
age <- html %>% html_nodes("#the40man .right:nth-child(8)") %>% html_text()
first_year <- html %>% html_nodes("#the40man .right:nth-child(14)") %>% html_text()

royals_2026 <- tibble(
  name = name,
  age = age,
  position = position,
  first_year = first_year
)
```


## Some issues with Web-Scraping

1. Data not always able to be scraped

```{r}
url <- "https://www.nhl.com/hurricanes/roster"
html <- read_html(url)
html %>% html_nodes("") %>% html_text()
```

- In many cases, this happens because you’re trying to scrape a website that dynamically generates the content of the page with javascript. This doesn’t currently work with `rvest`, because `rvest` downloads the raw HTML and doesn’t run any javascript.
  - There are work arounds for this but goes beyond the scope of this class..


## Working with Tables

If you’re lucky, your data will be already stored in an HTML table, and it’ll be a matter of just reading it from that table

```{r}
url <- "https://www.baseball-reference.com/teams/KCR/2026-roster.shtml"
html <- read_html(url)


html %>%
  html_element("table") %>%
  html_table()
```

## Your Turn 

1. Find a website that you may want to scrape (i.e. FDA, Census, NCAA Stats, Creighton Athletics, etc.) 
2. Read the URL into R and save the url as an object. 
3. Using the SelectorGadget on your website and look for a node with data that you want to collect.



